### 2011-04-26 Spatial_adaptation/local-adaptation.tex
/* Local Adaptation */
/* Local Adaptation */ latex oops
/* Local Adaptation */ small addition
oops equation

@@

This from ``NEQwiki, the nonlinear equations encyclopedia'', \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
We want to integrate
\[
    \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
        \sqrt{2}{s} \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } ,
\]
with $C=2K/s$.
Let $\phi^2(1-\phi/3)+C = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
and change variables first to $y^2=(\alpha-\phi$, 
and then to $x = y/\sqrt{\alpha-\beta}$, so that
\begin{align*}
    \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } 
        = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-p^2) } } \\
        = \frac{ - 2 dx }{ \sqrt{\alpha-\gamma} \sqrt{ (1-x^2) (1-S^2 x^2) } } ,
\end{align*}
with $S^2 = (\alpha-\beta)/(\alpha-\gamma)$.
Now Jacobi's ``incomplete elliptic integral of the first kind'' is
\[
    F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) } .
\]

>>>>

\newcommand{\sn}{\mathop{\mbox{sn}}}

@@

This from ``NEQwiki, the nonlinear equations encyclopedia'', \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
We want to integrate
\[
    \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
        \sqrt{2}{s} \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } ,
\]
with $C=2K/s$.
Let $\phi^2(1-\phi/3)+C = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
and change variables first to $y^2=(\alpha-\phi$, 
and then to $x = y/\sqrt{\alpha-\beta}$, so that
\begin{align*}
    \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } 
        = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-p^2) } } \\
        = \frac{ - 2 dx }{ \sqrt{\alpha-\gamma} \sqrt{ (1-x^2) (1-S^2 x^2) } } ,
\end{align*}
with $S^2 = (\alpha-\beta)/(\alpha-\gamma)$.
Now Jacobi's ``incomplete elliptic integral of the first kind'' is
\[
    F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) }} ,
\]
and the Jacobian elliptic function $\sn(x;k)$ is the inverse: $F(\sn(x;k);k) = x$.

As $k \to 0$, $\sn(x;k) \to \sin(x)$, while as $k \to 1$, $\sn(x;k) \to \sinh(x)$.

\section{Mutation rates}

Barton (Gen Res 87), has expressions for the probability of establishment as a function of spatial location in fixed landscape of advantageous pockets:
The probability of establishment beginning at state p is <latex>F(p) = 1 - \exp(-4Ns\int \phi(x) p(dx))</latex>,
where 
\[ \phi(x)^2 = L\phi(x) ; \qquad \mbox{and| \qquad \frac{\partial \phi}{\partial x} = 0 \qquad \mbox{as}\; x \to \infty .\]
Here $L\phi = \Delta\phi + f\phi$, where $$sf(x)$$ is the selective advantage is location $x$.
This form follows by assuming that the offspring of each individual go extinct independently (as in a branching process),
and so the probability of extinction is a product over the extant individuals;
there are $2N p(x) dx$ individuals at $x$
and the probability of exinction of a single individual at $x$ is $\exp (2s \phi(x))$;
Barton says this is derived by [http://www.bibsonomy.org/bibtex/23c3cf150aea743a3bfe52155d00cbf7b/peter.ralph Pollack] (J Appl Prob '66) using branching processes (and assuming Poisson offspring); 
Pollack actually does it in a discrete world; Barton is (silently) taking a continuum limit.
Barton also motivates by linearizing about establishment, and solves some interesting cases.

So, let $f(x) = s(x)/s_b$ be the rescaled selective benefit, and we want to solve
\[
    \phi''(x) = - f(x) \phi(x) + \phi(x)^2 .
\]
This fits into the framework above, if $f(x)$ is piecewise constant,
since
\[
    d \phi = \pm dx \sqrt{ K + f \phi + \phi^2 } .
\]
Barton gives the example we want...

### 2011-04-26 standing-patches-paper.tex
oops equation

Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
Therefore, if we define $\chi$ to be the unique positive solution to
\[
    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
\]
then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
before it is expected that one other successful mutation would have arisen in the area covered so far.
The explicit formula for $\chi$ is cumbersome but straightforward; here we omit the lengthy expression.
(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
This can be rewritten 
\begin{equation} \label{eqn:defines_chi}
    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^2 = (\pi \rho \mu)^{-1} .
\end{equation}
From this we see that $\chi$ increases linearly with $\sqrt{s_b}$,
and that the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
If $\sigma s_d \gg 1$, then standing variation is relatively unimportant, 
and this reduces to the characteristic length of \citet{ralphcoop2010}, $\chi_+ = (v/\lambda \pi)^{1/3}$.
On the other hand, if $\sigma s_d \ll 1$, then most adaptation come from standing variation, 
and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-2}$. 
% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.

>>>>

Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
Therefore, if we define $\chi$ to be the unique positive solution to
\[
    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
\]
then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
before it is expected that one other successful mutation would have arisen in the area covered so far.
The explicit formula for $\chi$ is cumbersome but straightforward; here we omit the lengthy expression.
(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
This can be rewritten 
\begin{equation} \label{eqn:defines_chi}
    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^3 = (\pi \rho \mu)^{-1} .
\end{equation}
From this we see that $\chi$ increases linearly with $1/\sqrt{s_b}$,
and that the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
If $\sigma s_d \gg 1$, then standing variation is relatively unimportant, 
and this reduces to the characteristic length of \citet{ralphcoop2010}, $\chi_+ = (v/\lambda \pi)^{1/3}$.
On the other hand, if $\sigma s_d \ll 1$, then most adaptation come from standing variation, 
and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-2}$. 
% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.
